Notes on Open Malaria

### 2018-12-14 documents/main.tex
updated

\section{What is Open Malria?}
Open malaria is for simulating malaria epidemology and the impacts on that epidemiology of interventions against malaria. 
 It is based on microsimulations of Plasmodium falciparum malaria in humans, originally developed for simulating malaria vaccines.
 These models simulate the dynamics of malaria parasitaemia in the course of an infection, of transmission, of immunity, and of the processes leading to illness and death. The system is set up to simulate malaria in village, or district size human populations

@@

\begin{itemize}
\item Endemic in 99 countries, 219 million cases, 660,000 deaths 
\item 90\% in Africa
\item Mortality rates fell by 26\% around the world in the past decade
\end{itemize}
\section{How can Open malaria help in the real world?}
\begin{figure}[h!]
\includegraphics[width=\textwidth]{images/research_process.png}
\caption{Where OM fits into the real world}
\end{figure}
\section{Simulation Options}

@@

\end{itemize}
\section{How to Use Open Malaria}

@@

\section{How does Malaria spread?}
\begin{itemize}
\item Deterministic model of malaria transmission based on the mosquito feeding cycle
\item Linked to models of interventions
\item Driven by EIR
\item For input EIR values, you may need to work backwards to determine pre-experiment level
\end{itemize}

\begin{figure}
\includegraphics[width=\textwidth]{images/malariacycle.png}
\caption{The mosquito feeding cycle. Malaria is spread through this cycle.}
\end{figure}

\section{Modelling different treatments}

\section{Running a simulation}

\text{Run ./openMalaria --scenario MyScenario.xml}
\section{Scenario transmission}

@@

\section{The phases of the simulator}
\begin{itemize}
\item the XML document is read, model constants are set and some variables initialised
\item an initial human population is created
\item a warm-up period of one human life span is run
\item there may be some extra simulations to adjust the mosquito population to the pre-specified Transmission situation.
\item the intervention phase starts.
\end{itemize}
During these different phases the feedback from infections in humans to those in mosquitoes is handled in different ways

@@

\section{Outputs}
The simulation results of the outputs you have chosen from Step #2 are produced in the form of two separate .txt files (ctsout.txt and output.txt) that can be imported into your program of choice for analysis (i.e. Excel, R, STATA, etc)
\section{Tom stuff}

@@

\section{Gunes Notes}
\begin{itemize}
	\item ABC - talk about this in the context to MCMMC
	\item The forward simulator represents some likelihood function that can be tractable or intractable. 
	\item Within forward simulators the randomness comes from rejection sampling statements , but conditioning within the simulators is hard
this is why we need to use the prob prog system . 
\end{itemize}

>>>>

\section{What is Open Malria?}
Open malaria is for simulating malaria epidemiology and the impacts on that epidemiology of interventions against malaria. 
 It is based on microsimulations of Plasmodium falciparum malaria in humans, originally developed for simulating malaria vaccines.
 These models simulate the dynamics of malaria parasitaemia in the course of an infection, of transmission, of immunity, and of the processes leading to illness and death. The system is set up to simulate malaria in village, or district size human populations

@@

\begin{itemize}
\item Endemic in 99 countries, 219 million cases, 660,000 deaths 
\item 90\% in Africa
\item Mortality rates fell by 26\% around the world in the past decade
\end{itemize}

\subsection{How does Malaria spread?}
\begin{itemize}
\item Deterministic model of malaria transmission based on the mosquito feeding cycle
\item Linked to models of interventions
\item Driven by EIR
\item For input EIR values, you may need to work backwards to determine pre-experiment level
\end{itemize}

\begin{figure}[ht!]
\includegraphics[width=\textwidth]{images/malariacycle.png}
\caption{The mosquito feeding cycle. Malaria is spread through this cycle.}
\end{figure}

\section{How can Open malaria help in the real world?}

It helps by enabling researchers to propose experiments, based on a number of observations that are collected on the field.

\begin{figure}[h!]
\includegraphics[width=\textwidth]{images/research_process.png}
\caption{Where OM fits into the real world}
\end{figure}
\section{Simulation Options}

@@

\end{itemize}

\subsection{Modelling different treatments}

\section{How to Use Open Malaria}

@@

\subsection{Scenario transmission}

@@

\subsection{Running a simulation}
Within the same directory structure we add the following files:

\begin{itemize}
\item densities.csv
\item scenario\_xx.xsd
\end{itemize}
Then there must exist at least on file <MyScenario.xml>

Then to run we use the following command:
\text{./openMalaria --scenario MyScenario.xml}

To ensure that it is possible to execute the openMalaria object, it must be within 
the execution path.

\subsection{The phases of the simulator}
\begin{itemize}
\item the XML document is read, model constants are set and some variables initialized
\item an initial human population is created
\item a warm-up period of one human life span is run
\item there may be some extra simulations to adjust the mosquito population to the pre-specified Transmission situation.
\item the intervention phase starts.
\end{itemize}
During these different phases the feedback from infections in humans to those in mosquitoes is handled in different ways

@@

\section{Outputs}
The simulation results of the outputs you have chosen from Step 2 are produced in the form of two separate .txt files (ctsout.txt and output.txt) that can be imported into your program of choice for analysis (i.e. Excel, R, STATA, etc)
\section{Tom stuff}

@@

\section{Gunes Notes}
\begin{itemize}
	\item ABC - talk about this in the context to MCMC
	\item The forward simulator represents some likelihood function that can be tractable or intractable. 
	\item Within forward simulators the randomness comes from rejection sampling statements , but conditioning within the simulators is hard
this is why we need to use the prob prog system . 
\end{itemize}
